Very simplified likelihood

### 2024-02-23 inst/very_simplified_likelihood.Rmd
more work

After simplifying and taking an asymptotic expansion we find

\[
\mathcal{S}=\log\mathcal{L}_{r,n}(a)=\log(B-1)-B\frac{r+1}{n} + \frac{B^2-B(5+r^2)/2}{n^2}
+ K + \mathcal{O}\left(n^{-3}\right)
\]

(where $K$ is an arbitrary constant).  If we then neglect terms of
order $n^{-2}$ and above, viz $\log(B-1)-B\frac{r+1}{n}$, observing
that $\partial^2\mathcal{S}/\partial B^2=-(B-1)^{-2}<0$, the evaluate
would be $\hat{a}=\frac{n}{n+r+1}$ [although this seems to be a poor
approximation to the exact evaluate]

@@

(denominator has $n+1$ terms).  In R:

@@

MLEa <- function(r,n,give=FALSE){
	out <- uniroot(f=function(a){fdash(a,r,n)},interval=c(0.01,0.99))
	if(!give){out <- out$root}
	return(out)
}

rbind(
c(MLE(1,9),MLEa(1,9)),
c(MLE(3,9),MLEa(3,9)),
c(MLE(3,7),MLEa(3,7))
)
```

```{r usestirling}
rbind(
c(MLE(10 , 100),100 / 111),
c(MLE(100,1000),1000/1101),
c(MLE(500,5000),5000/5501)
)

rbind(
c(MLEa(10 , 100),100 / 111),
c(MLEa(100,1000),1000/1101),
c(MLEa(500,5000),5000/5501)
)
```

@@

```{r,label=likevis,cache=F}
plotter <- function(...){
a <- seq(from=0,to=1,by=0.01)
n <- 7
jj <- f_vec_arn(a,3,n,log=FALSE)
plot(a,jj/max(jj,na.rm=TRUE),type='n')
grid()
for(r in seq(from=0,to=n)){
  y <- f_vec_a(a,r,n,log=FALSE)
  if(r==0){y[length(y)] <- 1}  
  y <- y/max(y,na.rm=TRUE)
  if(r>0){y[length(y)] <- 0}
  points(a,y,type='l',lwd=4,col=rainbow(n+1)[r+1])
  segments(x0=r/(n),y0=0.9,y1=1)
  text(x=r/(n),y=0.8,r)
}
}
plotter()
pdf(file="ninelikes.pdf")
plotter()
dev.off()
```

@@

```{r,label=fishbish,cache=TRUE}
x <- seq(from=0.1,to=0.95,by=0.01)
jj <- as.matrix(expand.grid(x,x))
L <- apply(jj,1,supp2_ed,rank[wp]-1,rank[wa]-1,class_size[wp]-1,class_size[wa]-1)
L <- L - max(L,na.rm=TRUE)
L <- matrix(L,length(x),length(x))
```

>>>>

Note that if we use instead $n~\sim n^ne^{-n}\sqrt{2\pi
n}\left(1+\frac{1}{12n}\right)$, then only terms of
$\mathcal{O}(n^{-2})$ and above are changed; the first-order terms and
indeed the zeroth order terms are unaffected.  Anyway, after
simplifying and taking an asymptotic expansion to order $n^{-1}$, we
find, to first order,

\[
\mathcal{S}=\log\mathcal{L}_{r,n}(a)=\log(B-1)-B\frac{r+1}{n} 
+ K + \mathcal{O}\left(n^{-2}\right)
\]

(where $K$ is an arbitrary constant).  Observing that
$\partial^2\mathcal{S}/\partial B^2=-(B-1)^{-2}<0$, the evaluate would
be unique; we have
$\hat{B}=\frac{n+r+1}{r+1}+\mathcal{O}\left(n^{-1}\right)$, or
alternatively $\hat{a}=\frac{n}{n+r+1}$.  I present some numerical
results below.

@@

(denominator has $n+1$ terms).  Observe that this agrees with the
$r<n$ case above (substituting $r=n$).  We may introduce some R idiom:

@@

MLEa <- function(r,n,give=FALSE,small=1e-4){
	out <- uniroot(f=function(a){fdash(a,r,n)},interval=c(small,1-small))
	if(!give){out <- out$root}
	return(out)
}

showdiff <- function(x,y){c(way1=x,way2=y,diff=x-y,logratio=log(x/y))}

rbind(
showdiff(MLE(1,9),MLEa(1,9)),
showdiff(MLE(3,9),MLEa(3,9)),
showdiff(MLE(3,7),MLEa(3,7))
)
```

```{r usestirling}
rbind(
showdiff(MLE(10, 100), 100 / 111),
showdiff(MLE(10, 500), 500 / 511),
showdiff(MLE(10,1000),1000 /1011),
showdiff(MLE(10,5000),5000 /5011)
)
```

```{r usestirlinga}
rbind(
showdiff(MLEa(10, 100), 100 / 111),
showdiff(MLEa(10, 500), 500 / 511),
showdiff(MLEa(10,1000),1000 /1011),
showdiff(MLEa(10,5000),5000 /5011)
)

```

@@

```{r,label=likevis,cache=F}
plotter <- function(...){
a <- seq(from=0,to=1,by=0.01)
n <- 7
jj <- f_vec_arn(a,3,n,log=FALSE)
plot(a,jj/max(jj,na.rm=TRUE),type='n')
grid()
for(r in seq(from=0,to=n)){
  y <- f_vec_a(a,r,n,log=FALSE)
  if(r==0){y[length(y)] <- 1}  
  y <- y/max(y,na.rm=TRUE)
  if(r>0){y[length(y)] <- 0}
  points(a,y,type='l',lwd=4,col=rainbow(n+1)[r+1])
}
}
plotter()
pdf(file="ninelikes.pdf")
plotter()
dev.off()
```

@@

```{r,label=fishbish,cache=TRUE}
x <- seq(from=0.1,to=0.99,by=0.01)
jj <- as.matrix(expand.grid(x,x))
L <- apply(jj,1,supp2_ed,rank[wp]-1,rank[wa]-1,class_size[wp]-1,class_size[wa]-1)
L <- L - max(L,na.rm=TRUE)
L <- matrix(L,length(x),length(x))
```
